Strong lensing selection effects

### 2022-08-03 paper/seleff.tex
large extended source

We explain how this definition applies in practice with a few examples.
%We can consider a few examples.
For simplicity, we focus on the case of a source with a single surface brightness peak, such as a S\'{e}rsic profile. All of the sources considered in this work belong to this family of surface brightness models.
We adopt the following procedure to determine the number of detected peaks.
Given the observed image of a lensed source, we define its footprint as the ensemble of pixels where the source is detected with $S/N>2$. The footprint is in general composed of multiple disconnected regions, corresponding to the different images. In order to only include images that can be clearly identified, we add the condition that the integrated $S/N$ of each separate region must be $S/N > 10$. This condition has the effect of removing from the source footprint any isolated region consisting only of a very small number of pixels. In a real-world application, it would be very hard to classify such marginal detections as images.
If, after applying this cut, the source footprint is spread over multiple separated regions, then the system is classified as a lens.
If the source footprint consists instead of a single region, we iteratively increase the surface brightness threshold used to define the footprint and count the number of isolated regions with $S/N>10$. The maximum number of detected regions defines the number of detected peaks.

@@

We can identify three different regimes in strong lensing of extended sources, depending on the relative size of the source and of the lens caustics.
In the limit of a very small source, the image configurations that are produced are qualitatively similar to those that can be obtained in the point-source case.
When the source and caustic size are comparable, the multiple images tend to be blended into arcs or rings. 
Most, if not all, of the galaxy-scale strong lenses known belong to these two categories.
However, there is a third regime in strong lensing, corresponding to the case in which the source size is much bigger than the caustics, such as in the fifth example of \Fref{fig:lensdef}.
In this regime, the overall size and total flux of the source are roughly preserved, and the lens produces only a relatively minor perturbation on a localised region of the image.
Strong lenses of this kind can be difficult to detect, especially if the region of the image subject to strong lensing overlaps with the light from the foreground galaxy. 
%For this reason, the best known examples of strong lensing in the large source size regime are produced by objects with a relatively low brightness that act as perturbers on images that are strongly lensed by a larger mass \citep[see e.g.][]{Veg++12}.
The ultimate limiting factor to the detection of strong lenses in the large source size regime, however, is the ability to spatially resolve the multiple images. This limit is set by the size of the point spread function (PSF).

@@

We compute $\crosssect$ in a series of different scenarios of increasing complexity.
%We first introduce the basic equations of gravitational lensing in section \ref{ssec:lensbasics}.
%Then, in section \ref{ssec:profile}, we present the model family adopted to describe the radial density profile of the lenses in our experiments.
The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments. We describe this in section \ref{ssec:profile}. %, among with the basic equations of gravitational lensing.
%In section \ref{ssec:axsymmpoint} we consider axisymmetric lenses and point sources.
In section \ref{ssec:axsymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
In section \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in section \ref{ssec:ellext} we replace point sources with extended sources.

%\subsection{Strong lensing basics}\label{ssec:lensbasics}

@@

%Lensing-related quantities depend on the projected mass distribution $\Sigma(R)$.
%For the purpose of computing lensing-related quantities, the projected massdistribution $\Sigma(R)$ is needed.
%For a S\'{e}rsic profile, this is given by
The projected surface mass density of a S\'{e}rsic profile is given by
\begin{equation}
\Sigma(R) = \Sigma_0 \exp{\left\{-b\left(\frac{R}{\reff}\right)^{1/4}\right\}},
\end{equation}
with
\begin{equation}
\Sigma_0 = \frac{\mstar b^{2n}}{2\pi n \reff^2 \Gamma(2n)},
\end{equation}
where $\mstar$ is the total mass, $\reff$ is the radius enclosing half of the total mass, $n$ is the S\'{e}rsic index, $\Gamma$ is the incomplete Gamma function, and $b$ is given by \citep{C+B99}
\begin{equation}
b(n) \approx 2n -\frac13 + \frac{4}{405n} + \frac{46}{25515n^2} + O(n^{-3}).
\end{equation}
Throughout this paper we indicate with $\teff$ the angular size of the half-light radius.

@@

\Fref{fig:kappa} shows the dimensionless surface mass density profile of S\'{e}rsic + gNFW models with various values of the inner dark matter density slope $\gammadm$ and of the fraction of projected dark matter mass within the half-light radius, $\fdm$.
All of the profiles have a S\'{e}rsic index $n=4$, a dark matter scale radius equal to ten times $\reff$, and are normalised in such a way that the Einstein radius is equal to the half-mass radius of the stellar component. 
Two main features emerge from \Fref{fig:kappa}. First, the baryons generally dominate the total density in the inner regions $(\theta < \tein)$, while the dark matter is the main component at large radii.
Second, models with different dark matter fractions and inner dark matter slopes can conspire to produce very similar total density profiles. 
%it is possible to build models with different dark matter fractions and inner dark matter slopes, but with very similar total density profiles. 
This is the case, for example, of the ($\fdm=0.5,\gammadm=1.5$) and the ($\fdm=0.2,\gammadm=2.0$) models (green and blue lines in \Fref{fig:kappa}).

@@

\Fref{fig:1dmag} shows the magnification of the secondary image as a function of source position. The secondary image is the one located in the region between the radial and tangential critical curves, opposite to the source with respect to the lens centre. In most practical cases this is the second brightest image.
As \Fref{fig:1dmag} shows, the magnification is very large for sources close to the lens (small values of $\beta$), decreases with increasing source position and then increases in close proximity to the radial caustic.
%In particular, for $\beta > 0.5\tein$, the magnification is smaller than unity, with the exception of a thin slice near the caustic.
While for the model with $f_{\mathrm{DM}}=0.8$ (purple curve) the magnification is above unity everywhere, other lens models can produce highly de-magnified secondary images for large values of $\beta$.
Depending on the intrinsic brightness of the lensed source, these images may or may not be detected.
%This has two implications for the lensing cross-section. First, a larger caustic size does not necessarily translate into a larger value of $\crosssect$. Second, the cross-section depends on the intrinsic brightness of the lensed source: an extremely bright point source can be detected regardless of its position in the source plane, but a source with a finite brightness can go undetected depending on its value of $\beta$.

@@

We can see a clear trend between $\Delta m$ and the lensing cross-section, in both panels of \Fref{fig:fixedap_cs}: $\crosssect$ is larger for brighter sources.
The trend saturates below a certain $\Delta m$, for sufficiently small values of $\gammadm$ or for sufficiently large values of $\fdm$.
%These are cases in which the source is so bright that, as long as it is located within the radial caustic, it always produces at least two detectable images. 
In these cases the lensing cross-section coincides with the full area enclosed within the caustic, and further increasing the source brightness does not result in an increased value of $\crosssect$.

@@

We measured $\crosssect$ for lenses with a fixed radial density profile and different ellipticities, with respect to point sources of diferent brightnesses.
%In particular, we took a lens at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.5$, $\reff=7$~kpc, $\log{\mdmfive}=11.0$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$. 
%In particular, we took a lens at redshift $z_{\mathrm{l}}=0.3$, with $\log{\mstar}=11.19$, $\reff=4.45$~kpc, $\fdm=0.5$, $r_s=100$~kpc, $\gammadm=1.5$, and a source at redshift $z_{\mathrm{s}}=1.5$. This corresponds to a lens with Einstein radius equal to the half-light radius, both equal to $1''$.
In particular, we set $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\tein=\teff$, and set the ellipticities of both the baryonic and dark matter components to be the same, with the same orientation of the major axis.
This is the lens model used to produce the caustics plot of \Fref{fig:ellcaust}.
That figure suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
Therefore we expect the strong lensing cross-section to be a weak function of ellipticity.

We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources over a given area that includes the caustic, then used {\sc Glafic} to solve the lens equation, find the number of images and their magnification. We then multiplied the area over which sources are located by the fraction of those that are strongly lensed according to the criterion of \Eref{eq:pointdet}.
%\Fref{fig:ellcaust} suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
%Therefore we expect the strong lensing cross-section to be a weak function of ellipticity.
The resulting $\crosssect$ is shown in \Fref{fig:ellpoint_cs} as a function of the minor-to-major axis ratio $q$.

@@

Most of the sources that result in detectable lenses produce two detectable images. These are sources that are located in the region enclosed bewteen the radial and the tangential caustic. 
%In most cases, only two of these three images are detected: the third one, close to the centre, is usually highly de-magnified.
If the source is located within the tangential caustic, however, four detectable images are usually created.
Lenses with four visible images, usually referred to as quad lenses, are sometimes given a high priority in certain lensing studies, because they offer more constraints compared to double lenses. For instance, quads make up the majority of the lenses used so far in time-delay studies \citep{Mil++20}.
For this reason we also computed an alternative lensing cross-section, in which the definition of a lensing event requires the detection of four images, instead of two.
This is plotted in \Fref{fig:ellpoint_cs} with dashed lines.
The cross-section for quads is a strong function of lens ellipticity, for bright sources.
This is a consequence of the fact that the area enclosed within the tangential caustic, which is where a source needs to be in order to produce four images, increases with increased lens ellipticity, as \Fref{fig:ellcaust} shows.
For sources that are intrinsically fainter than the detection limit, however, the quad cross-section is extremely small.

@@

In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the surface brightness distribution with its radial profile, shape and orientation, and the point spread function (PSF). 
For the sake of reducing the dimensionality of the problem, we neglected the PSF and focus on circularly symmetric sources. We also fixed the surface brightness profile to an exponential disk (i.e. a S\'{e}rsic model with $n=1$).
We then took a lens model with the same parameters used in section \ref{ssec:ellpoint} and a minor-to-major axis ratio of $q=0.7$.
We created a grid of pixels with $0.05''$ size, then computed the lensing cross-section by simulating a large number of images of extended sources with {\sc Glafic} and measuring the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
We carried out experiments with sources with different values of the total flux and varying half-light radius, $\theta_{\mathrm{e,s}}$.
The results are shown in \Fref{fig:extcs}.
The total flux of the source, indicated in the legend, is measured in units of the sky background fluctuation within an area equal to the square of the Einstein radius. We assumed that the background noise is an uncorrelated Gaussian field, for this purpose.

\begin{figure}
\includegraphics[width=\columnwidth]{ell_ext_cs.eps}
\caption{
Strong lensing cross-section as a function of source half-light radius.
The lens model is the same as in \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
Each solid line corresponds to a different value of the total flux of the source. The flux $f$ is expressed in terms of the background noise fluctuation measured over an area equal to $\tein^2$.
The vertical dashed lines correspond to the maximum size for which a galaxy with a given flux can be detected in the absence of lensing.
%Solid grey curves are lines of constant effective surface brightness, $f/(2\pi\theta_s^2)$.
\label{fig:extcs}
}
\end{figure}

@@

%The limit posed by lensing to the maximum detectable source size can be understood as follows.
%In the point source case, at a given flux, only sources that produce at least two images with a sufficiently high magnification are classified as strongly lensed.
%In most practical cases, the limiting factor is the magnification of the secondary image, $\mu_2$, which is a strong function of the source position (see \Fref{fig:1dmag}).
%Only sources that lie within a region of large $\mu_2$, the area of which is equal to the strong lensing cross-section, give rise to a strong lens configuration.
%When the source is extended, different parts of it are magnified by different amounts. If the source size is comparable to or larger than the region of high magnification, then its net magnification averaged over its extent will be lower than in the point-source case, and the secondary image will drop below the detection limit.
%This effect adds to the limit on the source surface brightness in the absence of lensing to produce the trend observed in \Fref{fig:extcs}.

%In the point source case, the necessary and sufficient condition for a source to be strongly lensed is that two detectable images are produced.
%The detection limit of the survey and the intrinsic flux of the source set a requirement on the minimum magnification that two images of the source must have in order for them to be detected.
%The limiting factor is set by the magnification of the secondary image: given a source flux, 
%In the point source case, at a given flux, only sources that lie within a certain region give rise to at least two detectable images, which are required to classify a system as a strong lens. 
%The extent of this region, which is equal to the strong lensing cross-section, is determined by the magnification required in order for the secondary image to be detected (see \Fref{fig:1dmag}).
%The area of this region is equal to the strong lensing cross-section, and decreases for fainter sources.

@@

To determine the density profile of the halo we relied on a theoretically motivated model that takes into account the effect of baryons on the dark matter.
We assumed the dark matter distribution to be initially described by an NFW profile with a concentration\footnote{The concentration is the ratio between the halo virial radius and the scale radius $r_s$.} of five, then used the prescription of \citet{Cau++20} to model the response of the halo to the infall of baryons. % on the basis of the present stellar mass distribution. 
In the \citet{Cau++20} model the halo response is approximated with an analytical function that depends on the present stellar mass distribution, and typically results in a more concentrated and steeper density profile compared to the original NFW model. 
Finally, we fitted a gNFW profile to the surface mass density of the contracted halo. By doing this, we were able to fully describe the dark matter density profile with three parameters: $\mhalo$, $\gammadm$ and $r_s$.

The dark matter density profile defined in this way is determined uniquely by the halo mass, the stellar mass and the half-light radius (the more concentrated the stellar distribution, the stronger the halo response, the steeper the dark matter density profile).
However, we wanted to explore scenarios in which there is scatter in the dark matter inner slope. Therefore, we applied a random scatter on $\gammadm$ of amplitude $\sigma_{\gamma}$. We explored different values for $\sigma_{\gamma}$, as explained in the next section.

>>>>

We explain how this definition applies in practice with a few examples.
For simplicity, we focus on the case of a source with a single surface brightness peak, such as a S\'{e}rsic profile. All of the sources considered in this work belong to this family of surface brightness models.
We adopt the following procedure to determine the number of detected peaks.
Given the observed image of a lensed source, we define its footprint as the ensemble of pixels where the source is detected with $S/N>2$. The footprint is in general composed of multiple disconnected regions, corresponding to the different images. In order to only include images that can be clearly identified, we add the condition that the integrated $S/N$ of each separate region must be $S/N > 10$. This condition has the effect of removing from the source footprint any isolated region consisting only of a very small number of pixels. In a real-world application, it would be very hard to classify such marginal detections as images.
If, after applying this cut, the source footprint is spread over multiple separated regions, then the system is classified as a lens.
If the source footprint consists instead of a single region, we iteratively increase the surface brightness threshold used to define the footprint and count the number of isolated regions with $S/N>10$. The maximum number of detected regions defines the number of detected peaks.

@@

We can identify three different regimes in strong lensing of extended sources, depending on the relative size of the source and of the lens caustics.
In the limit of a very small source, the image configurations that are produced are qualitatively similar to those that can be obtained in the point-source case.
When the source and caustic size are comparable, the multiple images tend to be blended into arcs or rings. 
Most, if not all, of the galaxy-scale strong lenses known belong to these two categories.
However, there is a third regime in strong lensing, corresponding to the case in which the source size is much bigger than the caustics, such as in the fifth example of \Fref{fig:lensdef}.
In this regime, the overall size and total flux of the source are roughly preserved, and the lens produces only a relatively minor perturbation on a localised region of the image.
Strong lenses of this kind can be difficult to detect, especially if the region of the image subject to strong lensing overlaps with the light from the foreground galaxy. 
The ultimate limiting factor to the detection of strong lenses in the large source size regime, however, is the ability to spatially resolve the multiple images. This limit is set by the size of the point spread function (PSF).

@@

We compute $\crosssect$ in a series of different scenarios of increasing complexity.
The model family adopted to describe the radial density profile of the lenses is the same in all of our experiments. We describe this in section \ref{ssec:profile}.
In section \ref{ssec:axsymmpoint} we show calculations of the strong lensing cross-section in the case of axisymmetric lenses and point sources.
In section \ref{ssec:ellpoint} we generalise the lens geometry to elliptical, while in section \ref{ssec:ellext} we replace point sources with extended sources.

@@

The projected surface mass density of a S\'{e}rsic profile is given by
\begin{equation}
\Sigma(R) = \Sigma_0 \exp{\left\{-b\left(\frac{R}{\reff}\right)^{1/4}\right\}},
\end{equation}
with
\begin{equation}
\Sigma_0 = \frac{\mstar b^{2n}}{2\pi n \reff^2 \Gamma(2n)},
\end{equation}
where $\mstar$ is the total mass, $\reff$ is the radius enclosing half of the total mass, $n$ is the S\'{e}rsic index, $\Gamma$ is the incomplete Gamma function, and $b$ is given by \citep{C+B99}
\begin{equation}
b(n) \approx 2n -\frac13 + \frac{4}{405n} + \frac{46}{25515n^2} + O(n^{-3}).
\end{equation}
Throughout this paper we indicate with $\teff$ the angular size of the half-light radius.

@@

\Fref{fig:kappa} shows the dimensionless surface mass density profile of S\'{e}rsic + gNFW models with various values of the inner dark matter density slope $\gammadm$ and of the fraction of projected dark matter mass within the half-light radius, $\fdm$.
All of the profiles have a S\'{e}rsic index $n=4$, a dark matter scale radius equal to ten times $\reff$, and are normalised in such a way that the Einstein radius is equal to the half-mass radius of the stellar component. 
Two main features emerge from \Fref{fig:kappa}. First, the baryons generally dominate the total density in the inner regions $(\theta < \tein)$, while the dark matter is the main component at large radii.
Second, models with different dark matter fractions and inner dark matter slopes can conspire to produce very similar total density profiles. 
This is the case, for example, of the ($\fdm=0.5,\gammadm=1.5$) and the ($\fdm=0.2,\gammadm=2.0$) models (green and blue lines in \Fref{fig:kappa}).

@@

\Fref{fig:1dmag} shows the magnification of the secondary image as a function of source position. The secondary image is the one located in the region between the radial and tangential critical curves, opposite to the source with respect to the lens centre. In most practical cases this is the second brightest image.
As \Fref{fig:1dmag} shows, the magnification is very large for sources close to the lens (small values of $\beta$), decreases with increasing source position and then increases in close proximity to the radial caustic.
While for the model with $f_{\mathrm{DM}}=0.8$ (purple curve) the magnification is above unity everywhere, other lens models can produce highly de-magnified secondary images for large values of $\beta$.
Depending on the intrinsic brightness of the lensed source, these images may or may not be detected.
%This has two implications for the lensing cross-section. First, a larger caustic size does not necessarily translate into a larger value of $\crosssect$. Second, the cross-section depends on the intrinsic brightness of the lensed source: an extremely bright point source can be detected regardless of its position in the source plane, but a source with a finite brightness can go undetected depending on its value of $\beta$.

@@

We can see a clear trend between $\Delta m$ and the lensing cross-section, in both panels of \Fref{fig:fixedap_cs}: $\crosssect$ is larger for brighter sources.
The trend saturates below a certain $\Delta m$, for sufficiently small values of $\gammadm$ or for sufficiently large values of $\fdm$.
In these cases the lensing cross-section coincides with the full area enclosed within the caustic, and further increasing the source brightness does not result in an increased value of $\crosssect$.

@@

We measured $\crosssect$ for lenses with a fixed radial density profile and different ellipticities, with respect to point sources of diferent brightnesses.
In particular, we set $\fdm=0.5$, $\gammadm=1.5$, $r_s=10\reff$, $\tein=\teff$, and set the ellipticities of both the baryonic and dark matter components to be the same, with the same orientation of the major axis.
This is the lens model used to produce the caustics plot of \Fref{fig:ellcaust}.
\Fref{fig:ellcaust} suggests that the size of the source plane area subject to strong lensing, the area enclosed within the outermost caustic, does not vary strongly with the ellipticity of the lens.
Therefore we expect the strong lensing cross-section to be a weak function of ellipticity.

We carried out the computation of $\crosssect$ by means of simulation: we generated a large number of point sources over a given area that includes the caustic, then used {\sc Glafic} to solve the lens equation, find the number of images and their magnification. We then measured the fraction of sources that are strongly lensed according to the criterion of \Eref{eq:pointdet} and multiplied this value by the area over which sources are located.
The resulting $\crosssect$ is shown in \Fref{fig:ellpoint_cs} as a function of the minor-to-major axis ratio $q$.

@@

Most of the sources that result in detectable lenses produce two detectable images. These are sources that are located in the region enclosed bewteen the radial and the tangential caustic. 
If the source is located within the tangential caustic, however, four detectable images are usually created.
Lenses with four visible images, usually referred to as quad lenses, are sometimes given a high priority in certain lensing studies, because they offer more constraints compared to double lenses. For instance, quads make up the majority of the lenses used so far in time-delay studies \citep{Mil++20}.
For this reason we also computed an alternative lensing cross-section, in which the definition of a lensing event requires the detection of four images, instead of two.
This is plotted in \Fref{fig:ellpoint_cs} with dashed lines.
The cross-section for quads is a strong function of lens ellipticity, for bright sources.
This is a consequence of the fact that the area enclosed within the tangential caustic, which is where a source needs to be in order to produce four images, increases with increased lens ellipticity, as \Fref{fig:ellcaust} shows.
For sources that are intrinsically fainter than the detection limit, however, the quad cross-section is extremely small, regardless of ellipticity.

@@

In the case of an extended source, the complexity of the problem is increased due to the addition of a series of features: the source surface brightness distribution, with its radial profile, shape and orientation, and the point spread function (PSF). 
Moreover, as we discussed in section \ref{ssec:lensdefext}, there are different regimes in strong lensing of extended sources, depending on the relative size of the source and the caustics of the lens.
For this reason, we split our analysis in two parts: first we explore the small source size regime, then we consider cases in which the source size is comparable to or bigger than the lens caustics.

\subsubsection{Small source sizes}

For the sake of reducing the dimensionality of the analysis, we focused on circularly symmetric sources. We also fixed the surface brightness profile to an exponential disk (i.e. a S\'{e}rsic model with $n=1$).
We then took a lens model with the same parameters used in section \ref{ssec:ellpoint} and a minor-to-major axis ratio of $q=0.7$.
We simulated a large number of images of extended sources with {\sc Glafic} and measured the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
For the small source size experiment, we used pixels with a size equal to $1/20\tein$ and convolved the images with a Moffat PSF with a FWHM of two pixels.
We also assumed that the background noise is an uncorrelated Gaussian field.
We carried out experiments with sources with different values of the total flux, $f$, and half-light radius, $\theta_{\mathrm{e,s}}$.
The results are shown in \Fref{fig:extcs}.
The total flux of the source, indicated in the legend, is measured in units of the sky background fluctuation within an area equal to the square of the Einstein radius.

\begin{figure}
\includegraphics[width=\columnwidth]{ell_ext_cs.eps}
\caption{
Strong lensing cross-section as a function of source half-light radius.
Small source size regime.
The lens model is the same as in \Fref{fig:ellpoint_cs}, with axis ratio $q=0.7$. The source is a circular exponential profile.
Each solid line corresponds to a different value of the total flux of the source. The flux $f$ is expressed in terms of the background noise fluctuation measured over an area equal to $\tein^2$.
The vertical dashed lines correspond to the maximum size for which a galaxy with a given flux can be detected in the absence of lensing.
The horizontal dotted line indicates the cross-section for a very bright point source (i.e. the area enclosed within the caustics).
%Solid grey curves are lines of constant effective surface brightness, $f/(2\pi\theta_s^2)$.
\label{fig:extcs}
}
\end{figure}

%We created a grid of pixels with $0.05''$ size, then computed the lensing cross-section by simulating a large number of images of extended sources with {\sc Glafic} and measuring the fraction of them that results in a strong lens, according to the definition of section \ref{ssec:lensdefext}.
%We carried out experiments with sources with different values of the total flux and varying half-light radius, $\theta_{\mathrm{e,s}}$.
%The results are shown in \Fref{fig:extcs}.
%The total flux of the source, indicated in the legend, is measured in units of the sky background fluctuation within an area equal to the square of the Einstein radius. We assumed that the background noise is an uncorrelated Gaussian field, for this purpose.

@@

\subsubsection{Large source sizes}

For the large source size case we fixed the surface brightness profile of the source and varied the Einstein radius of the lens.
In particular, we set the source half-light radius to $20$ pixels and adjusted its total flux in such a way that the $2\sigma$ detection footprint in the absence of lensing extends out to the half-light radius.
Then, starting from the lens model used in the previous section, we progressively increased the critical surface mass density to reduce the Einstein radius down to values comparable to the pixel size.
\Fref{fig:largesourcecs} shows the resulting strong lensing cross-section as a function of the ratio between lens Einstein radius and source half-light radius.

@@

To determine the density profile of the halo we relied on a theoretically motivated model that takes into account the effect of baryons on the dark matter.
We assumed the dark matter distribution to be initially described by an NFW profile with a concentration\footnote{The concentration is the ratio between the halo virial radius and the scale radius $r_s$.} of five, then used the prescription of \citet{Cau++20} to model the response of the halo to the infall of baryons. % on the basis of the present stellar mass distribution. 
In the \citet{Cau++20} model the halo response is approximated with an analytical function that depends on the present stellar mass distribution, and typically results in a more concentrated and steeper density profile compared to the original NFW model. 
Finally, we fitted a gNFW profile to the surface mass density of the contracted halo. By doing this, we were able to fully describe the dark matter density profile with three parameters: $\mhalo$, $\gammadm$ and $r_s$.
%
The dark matter density profile defined in this way is determined uniquely by the halo mass, the stellar mass and the half-light radius (the more concentrated the stellar distribution, the stronger the halo response, the steeper the dark matter density profile).
%However, we wanted to explore scenarios in which there is scatter in the dark matter inner slope. Therefore, we applied a random scatter on $\gammadm$ of amplitude $\sigma_{\gamma}$. We explored different values for $\sigma_{\gamma}$, as explained in the next section.
